---
title: "Final Project"
author: "Bridget Wojcik"
date: "2023-11-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Actors Data Set
**Introduce your data using complete sentences.**  What kind of data set are you looking at? Why did you choose this data set?

We chose to look at a data set of a sample of famous actor and look at their total gross, number of movies, average per movie, #1 movie, and the gross for that movie. We thought this would be an interesting topic because these famous actors are relevant to our age and generation. We liked this specific data set because it was already tidy, along with descending order of total gross which was very helpful. We want to further explore this data set whilst answering some research question that we formed. 

#load the data set
```{r actors}
#Load the correct packages to open the data set
library(dyplr) 
library(readr)

#read the csv file (Insert your own name and location of the data set) 
Actors <- read_csv ("/Users/bridgetwojcik/Downloads/actors.csv")
```

## Research Question 1

Does who the Actor is effect what their total gross is?

```{r pressure, echo=FALSE}
# Load necessary libraries
library(ggplot2)
library(stringr)  

ggplot(Actors, aes(x = str_wrap(Actor, width = 10), y = `Total Gross`)) +
  geom_point() +  # Create a scatterplot
  labs(title = "Scatterplot of Actor by Total Gross",  # Set plot title
       x = "Actor",  # Set x-axis label
       y = "Total Gross") +  # Set y-axis label
  theme(axis.text.x = element_text(angle = 75, hjust = 1, size = 4))  # Adjust x-axis text angle and size
```

#EDA2
```{r, echo= FALSE}
# Create a bar plot
ggplot(Actors, aes(x = str_wrap(Actor, width = 10), y = `Total Gross`)) +
  geom_bar(stat = "identity", fill = "red", color = "black") +
  labs(title = "Bar Plot of Actor by Total Gross",
       x = "Actor",
       y = "Total Gross") +
  theme(axis.text.x = element_text(angle = 75, hjust = 1, size = 5))


```
#Visual Table 

```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)



# Convert "Actor" to character
Actors$Actor <- as.character(Actors$Actor)

# Calculate overall mean, median, and mode for Total Gross
overall_mean <- mean(Actors$`Total Gross`)
overall_median <- median(Actors$`Total Gross`)
overall_mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

# Create a summary table
summary_table <- data.frame(
  Measure = c("Mean", "Median", "Mode"),
  Value = c(overall_mean, overall_median, overall_mode(Actors$`Total Gross`))
)

print(summary_table)


```


##Research Question 2 

Is there a notable difference between an actor's overall earnings (total gross) and the quantity of movies they have starred in? 

```{r}
# Convert "actor" to character
Actors$Actor <- as.character(Actors$Actor)
Actors$Num_Mov <- as.numeric(Actors$`Number of Movies`)


# Create a bar plot
ggplot(Actors, aes(x = Actor, y = Num_Mov)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Number of Movies by Actor",
       x = "Actor",
       y = "Number of Movies")

# Load necessary libraries
library(ggplot2)
library(stringr)  


# Load necessary libraries
library(ggplot2)
library(stringr)  

ggplot(Actors, aes(x = str_wrap(Actor, width = 10), y = `Total Gross`)) +
  geom_point() +  # Create a scatterplot
  labs(title = "Scatterplot of Actor by Total Gross",  # Set plot title
       x = "Actor",  # Set x-axis label
       y = "Total Gross") +  # Set y-axis label
  theme(axis.text.x = element_text(angle = 75, hjust = 1, size = 4))  # Adjust x-axis text angle and size
        
```
Shown here, is a new visualization which is a bar graph to compare the actor with their number of movies in the data set. We thought it was important to compare both graphs to look at how each actors number of movies compares to what their total gross is. After looking at these two graphs we can see that the number of movies do not necessarily align with what their total gross is. Robert DeNiro has the highest number of movies in this sample, but a lower total gross. Similarly, Anthony Daniels has the least amount of movies in this sample, but has the fifteenth highest total gross. 

```{r}
# Load necessary libraries
library(ggplot2)


# Convert "Actor" to character
Actors$Actor <- as.character(Actors$Actor)

# Set larger plot size
options(repr.plot.width = 8, repr.plot.height = 6)

# Create a scatter plot with Actor, Total Gross, and Number of Movies
ggplot(Actors, aes(x = `Number of Movies`, y = `Total Gross`, color = Actor)) +
  geom_point(size = 2) +
  scale_color_viridis_d() +
  labs(title = "Actor Statistics",
       x = "Number of Movies",
       y = "Total Gross",
       color = "Actor") +
  theme_minimal() +
  theme(axis.text = element_text(size = 3),
        legend.text = element_text(size = 3),
        legend.key.size = unit(0.5, "cm")) +
  xlim(c(0, max(Actors$`Number of Movies`))) +
  ylim(c(0, max(Actors$`Total Gross`)))
```

Here, is a graph that shows all of the variables we wanted to look at. It has the number of movies on the x-axis vs total gross on the y-axis and the points are color-coded to match with each individual actor. I really like this graph because it is more obvious to predict some kind of trend here. From what I see, this scatter-plot looks to have a positive linear correlation, so it simpler terms, as the number of movies increase, the total gross is increasing. This is a surprising trend to see given our predictions from before and how they don't correlate. 

```{r}
library(gt)


# Convert "Actor" to character
Actors$Actor <- as.character(Actors$Actor)

# Create a gt table with "Total Gross" and "Number of Movies" columns
actors_table <- gt(Actors[, c("Actor", "Number of Movies", "Total Gross")]) %>%
  fmt_number(
    columns = vars(`Number of Movies`, `Total Gross`),
    decimals = 0
  ) %>%
  fmt_currency(
    columns = vars(`Total Gross`),
    currency = "USD",
    decimals = 0
  ) %>%
  tab_header(
    title = "Actor",
    subtitle = "Table showing statistics for each actor."
  )

# Print the table
actors_table




```
This table is essentially a simpler way to read the data set because it now just includes the numbers we want to see, rather than all the other variables we are not looking at. You can also see a little easier how these numbers dont necessarily align. The number of movies does not seem to effect the total gross.

## Research Question 3
```{r, echo=FALSE}
# Load necessary libraries
library(readxl)
library(knitr)

# Load the dataset from Excel
Athletes <- read_excel("/Users/bridgetwojcik/Downloads/Athletes.xlsx")  # Adjust the path accordingly

```
In this plot here you can see that almost all of these athletes have around the same pay besides for a few outliers. Comparing this to the actors data, there appears to be less variety in the amount of pay/total gross. 

```{r, echo=FALSE}
# Create a dot plot 
dotchart(Athletes$Pay, labels = Athletes$Athlete,
         main = "Athlete Pay",
         xlab = "Pay",
         ylab = "Athlete",
         color = rainbow(length(Athletes$Athlete)),
         cex = 0.4)  # Adjust the font size 

# Add grid lines
grid()


```
Shown here is a dot chart, this is just another visualization to look at for the same variables. This time on this graph, the athletes are shown on the y-axis rather than the x. This is just another way to look at it differently. From looking here, however, it seems to be less evident that they all have a similar pay besides some outliers. Here it looks extremely skewed, which is something to think about because in our actors data, the total gross also seemed to be sort of all over the place. 

```{r, echo= FALSE}
# Create a data frame 
athlete_data <- data.frame(Athlete = Athletes$Athlete, Pay = Athletes$Pay)

# Print the data frame
print(athlete_data)

```
Here, is just a simple table to look at each athletes pay more clearly and really get a look at all of the numbers. Once you look at the numbers this way, it more evident to see that the pay decreases pretty consistently and the biggest jump is from Michael Jordan to Tiger Woods. 

##Conclusion
After making visualizations for each research questions along with tables, we got to explore more into actors total gross along with athlethes pay and were able to compare the two. The analysis suggests that predicting an actor's total gross is a complex task, and the number of movies alone does not provide a straightforward explanation. The comparison with highest paid athletes highlights the variability in earnings across different industries. Overall, the entertainment industry exhibits a diverse range of earnings, and multiple factors may contribute to an individual's success.
##Code Apendix
```{r}
library(dplyr)
library(readr)


Actors <- read_csv("/Users/bridgetwojcik/Downloads/actors.csv")

# Load necessary libraries
library(ggplot2)
library(stringr)  

ggplot(Actors, aes(x = str_wrap(Actor, width = 10), y = `Total Gross`)) +
  geom_point() +  # Create a scatterplot
  labs(title = "Scatterplot of Actor by Total Gross",  # Set plot title
       x = "Actor",  # Set x-axis label
       y = "Total Gross") +  # Set y-axis label
  theme(axis.text.x = element_text(angle = 75, hjust = 1, size = 4))  # Adjust x-axis text angle and size

# Create a bar plot
ggplot(Actors, aes(x = str_wrap(Actor, width = 10), y = `Total Gross`)) +
  geom_bar(stat = "identity", fill = "red", color = "black") +
  labs(title = "Bar Plot of Actor by Total Gross",
       x = "Actor",
       y = "Total Gross") +
  theme(axis.text.x = element_text(angle = 75, hjust = 1, size = 5))

# Load necessary libraries
library(ggplot2)
library(dplyr)



# Convert "Actor" to character
Actors$Actor <- as.character(Actors$Actor)

# Calculate overall mean, median, and mode for Total Gross
overall_mean <- mean(Actors$`Total Gross`)
overall_median <- median(Actors$`Total Gross`)
overall_mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

# Create a summary table
summary_table <- data.frame(
  Measure = c("Mean", "Median", "Mode"),
  Value = c(overall_mean, overall_median, overall_mode(Actors$`Total Gross`))
)

print(summary_table)

# Convert "actor" to character
Actors$Actor <- as.character(Actors$Actor)
Actors$Num_Mov <- as.numeric(Actors$`Number of Movies`)


# Create a bar plot
ggplot(Actors, aes(x = Actor, y = Num_Mov)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Number of Movies by Actor",
       x = "Actor",
       y = "Number of Movies")

# Load necessary libraries
library(ggplot2)
library(stringr)  


# Load necessary libraries
library(ggplot2)
library(stringr)  

ggplot(Actors, aes(x = str_wrap(Actor, width = 10), y = `Total Gross`)) +
  geom_point() +  # Create a scatterplot
  labs(title = "Scatterplot of Actor by Total Gross",  # Set plot title
       x = "Actor",  # Set x-axis label
       y = "Total Gross") +  # Set y-axis label
  theme(axis.text.x = element_text(angle = 75, hjust = 1, size = 4))  # Adjust x-axis text angle and size

# Load necessary libraries
library(ggplot2)


# Convert "Actor" to character
Actors$Actor <- as.character(Actors$Actor)

# Set larger plot size
options(repr.plot.width = 8, repr.plot.height = 6)

# Create a scatter plot with Actor, Total Gross, and Number of Movies
ggplot(Actors, aes(x = `Number of Movies`, y = `Total Gross`, color = Actor)) +
  geom_point(size = 2) +
  scale_color_viridis_d() +
  labs(title = "Actor Statistics",
       x = "Number of Movies",
       y = "Total Gross",
       color = "Actor") +
  theme_minimal() +
  theme(axis.text = element_text(size = 3),
        legend.text = element_text(size = 3),
        legend.key.size = unit(0.5, "cm")) +
  xlim(c(0, max(Actors$`Number of Movies`))) +
  ylim(c(0, max(Actors$`Total Gross`)))

# Load the gt package 
library(gt)

# Convert "Actor" column to character type
Actors$Actor <- as.character(Actors$Actor)

# Create a simple gt table with selected columns
actors_table <- gt(Actors[, c("Actor", "Number of Movies", "Total Gross")]) %>%
  
  # Make the selected numeric columns to have no decimals
  fmt_number(
    columns = vars(`Number of Movies`, `Total Gross`),
    decimals = 0
  ) %>%
  
  # Make the "Total Gross" column as currency in USD with no decimals
  fmt_currency(
    columns = vars(`Total Gross`),
    currency = "USD",
    decimals = 0
  )

# Print the final table
actors_table

# Create a horizontal bar plot 
barplot(height = Athletes$Pay, names.arg = Athletes$Athlete,
        main = "Athlete Pay",
        xlab = "Athlete",
        ylab = "Pay",
        col = rainbow(length(Athletes$Athlete)),
        cex.names = 0.6,  # Adjust the font size 
        las = 2)  # Rotate the labels

# Add grid lines
grid()

# Create a dot plot 
dotchart(Athletes$Pay, labels = Athletes$Athlete,
         main = "Athlete Pay",
         xlab = "Pay",
         ylab = "Athlete",
         color = rainbow(length(Athletes$Athlete)),
         cex = 0.4)  # Adjust the font size 

# Add grid lines
grid()

# Create a data frame 
athlete_data <- data.frame(Athlete = Athletes$Athlete, Pay = Athletes$Pay)

# Print the data frame
print(athlete_data)


```
This is the link to our actors data set in our github. We go the data set from a previous class and found the sourse at github.
https://github.com/Stat184-Fall2023/FP_BridgetW_VaibhavC_EricL/blob/main/actors.csv 

This is the link to the athletes data set. I then tidied the data into excel.
https://www.statista.com/statistics/1370671/highest-paid-athletes-ever/




